
```{r appendix, echo=FALSE}
torontodown::here_appendix()
```

# The First Appendix


```{r message=FALSE, eval=TRUE}
library(here)
library(kableExtra)
library(ggpubr)
library(tibble)
library(readxl)
library(openxlsx)
library(r4ss)
library(readr)
library(flextable)
library(officer)
library(r4ss)
```

\newpage

## Abstract 

Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is critical for achieving sustainable and ecosystem-based management fisheries in the Western Antarctic Peninsula. In this study, we implemented an integrated length-to-age model that not only incorporates diverse data sources related to krill population dynamics, but also enables the assessment of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status. We explored four scenarios that either accounted for or ignored ecosystem influences on krill population dynamics, yielding results consistent with the expected behavior of a stock assessment framework tailored to this type of species. Preliminary performance analyses reveal systematic differences across scenarios, highlighting increased bias in estimates of fishing mortality (F) and spawning stock biomass (SSB) in models that include ecosystem variables. Conversely, the inclusion of environmental and predator data results in lower estimates of recruitment, total biomass, and spawning biomass, offering a more precautionary—and arguably more realistic—representation of krill population dynamics. Despite ongoing challenges in incorporating ecosystem variables into integrated models like this, we argue that such factors should be embedded directly into the state equations that govern population dynamics, rather than addressed through pre and post analyses. Overlooking these drivers may lead to less accurate conclusions, hence limiting the effectiveness of management recommendations. Integrated modeling approaches that draw on extensive biological, ecological, fishery, and environmental datasets provide a robust foundation for developing management strategies aligned with CCAMLR’s objectives for conservation and sustainable use. Such approaches are particularly relevant for ecologically complex systems like the krill fishery in the Western Antarctic Peninsula.




*Keywords: Antarctic krill, population dynamics, length-to-age model, stock-recruitment, ecosystem variables, performance model, fisheries management, CCAMLR*

\newpage

## Introduction

Antarctic krill (*Euphausia superba*, hereafter krill) is one of the largest populations on the planet in terms of biomass and also a keystone species in the Antarctic ecosystem [@Atkinson2019a; @McBride2021; @Kawaguchi2024]. Currently, the population is predominantly concentrated in the Western Antarctic Peninsula (WAP), being the primary area for krill harvesting over the past four decades and also an area that is also a critical feeding ground for predators such as penguins, seals, and whales [@Watters2013; @Hill2016; @Gallagher2023a]. At the same time, the WAP is one of the regions most affected by climate change, experiencing rapid increases in ocean temperature, sea ice loss, and changes in primary productivity [@Turner2005; @Lima2013; @Atkinson2019a; @Morley2020; @Carrasco2021; @McBride2021]. These overlapping pressures have intensified scientific interest in understanding krill population dynamics from both an ecological and fisheries perspective in such a complex and changing environment [@Warwick-Evans2022; @Kawaguchi2024].

Given the importance of krill to the ecosystem and its commercial value, effectively managing krill stocks is essential to maintaining the balance of the Antarctic ecosystem while supporting sustainable fisheries. Recognizing this, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) has been working on a revised ecosystem-based approach [@Constable2011; @Watters2013; @Cavanagh2016; @Constable2023]. This strategy seeks to balance commercial harvesting with a precautionary framework that safeguards krill-dependent species in space and temporal scales. It comprises three key components: (i) stock assessments to establish precautionary harvest rates, (ii) periodic updates of krill biomass estimates at multiple spatial scales, and (iii) a spatial overlap analysis, to guide the spatial allocation of catches and mitigate ecosystem-wide risks [@Warwick-Evans2022; @Kawaguchi2024].

To provide scientific advice and recommendations on the status and standing stock of krill, CCAMLR has historically relied on dedicated scientific surveys [@Watters2013; @Hill2019; @Dornam2021; @Green2023]. However, changes in survey design—both methodological and spatial—have raised concerns about the reliability of these estimates, while the high costs associated with such surveys have further motivated the search for alternative stock estimation methods. More recently, and as part of a new strategy, CCAMLR has explored a variety of model-based (indirect) approaches to assess krill stock status and support fisheries management decisions [@Hill2019]. Some of these stock assessment models have used a single source of data—either fishery-dependent catches [@Wang2021] or survey-based monitoring programs [@Kinzey2011; @Kinzey2019; @Green2023]. Nowadays, the Generalized Yield Model (GRYM) [@Maschette2020] is a model-based approach used by CCAMLR to generate management recommendations, allowing adjustments to the decision rules to predefined catch limits set by CCAMLR [@Constable2000; @CCAMLR2010; @CCAMLR2021a]. However, model-based approaches exhibit limitations, including an effective integration of multiple data sources, incorporation of ecosystem-level variables into population dynamics, and accounting for the high spatial and temporal intrinsic variability of krill productivity [@Meyer2020; @Trathan2021].

Increasing evidence demonstrates that environmental conditions—such as sea surface temperature, sea ice extent, and phytoplankton availability—strongly influence krill recruitment, survival, and spatial distribution [@Murphy2007; @Flores2012; @Saba2014; @Morley2020; @Perry2020; @Walsh2020; @Vorrath2020; @Veytia2021]. Similarly, fluctuations in predator abundance and distribution can exert significant top-down control on krill populations [@Moosa2017; @Reisinger2022; @Baines2022; @Riaz2023; @Biuw2024]. There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and explicitly account for ecosystem interactions [@Marshall2019]. Neglecting these ecosystem drivers can lead to biased estimates of stock status and suboptimal management advice [@Haltuch2009; @Richards2016; @Marshall2019; @Crone2019]. Therefore, stock assessment frameworks aligned with CCAMLR’s new strategy should explicitly incorporate these ecosystem components to more accurately represent krill population processes.

Integrating ecological complexity into krill stock assessments presents multiple challenges. Long-term data collection is often affected by changes in monitoring programs and sampling methodologies [@Hill2019], while some key regions of the WAP lack sufficient data to construct reliable krill population indices. Additional difficulties arise from indirect drivers—such as environmental variability and predator-prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [@Constable2023]. Despite these challenges, substantial data exist on krill and associated environments. Fishery monitoring programs provide detailed biological data from catches through onboard scientific observers. Ecosystem monitoring, like the CCAMLR Ecosystem Monitoring Program (CEMP), tracks predator populations and environmental variability at key sites. Together, these initiatives support ecosystem-based management by offering insights into both direct fishery impacts and the broader ecological context influencing krill dynamics.

This diversity of data creates an opportunity to develop integrated assessment frameworks that account for the coupled biological, ecological, and fishery processes shaping krill populations. Spatially explicit approaches include the State-Space Assessment Model (SAM) [@Nielsen2021] and the Globally Applicable Area-Disaggregated General Ecosystem Toolbox (Gadget) [@Bartolino2011]. For statistical catch-at-age assessments, widely used frameworks include the C++ Algorithmic Stock Assessment Laboratory (CASAL) [@Bull2004] and Stock Synthesis (SS3) [@Methot2013], among others. These advanced tools offer robust methodologies to address existing data gaps while incorporating the full complexity of ecosystem and population dynamics [@Hoyle2020].

This study aims to integrate various data sources on krill (*Euphausia superba*), including fishery operations (e.g. catches and effort distribution), biological characteristics (e.g. growth rates and reproductive parameters), spatiotemporal population dynamics, and key ecosystem drivers (e.g. environmental forcing and predator-prey interactions) into a unified stock assessment framework. Specifically, we use an integrated length-to-age model implemented in Stock Synthesis (SS3), which enables the explicit inclusion of environmental covariates affecting biomass, recruitment patterns, and productivity, as well as the potential impacts of predator populations in natural mortality configuration. This approach strengthens the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR's ecosystem-based management approach [@Constable2011; @Cavanagh2016; @Hill2019; @Constable2023; @Zhao2023; @CCAMLRWGSAM2024; @WGEMM2024]. The primary objective is to evaluate the statistical performance of alternative model configurations representing different hypotheses about ecosystem influences using a suite of diagnostic tools and performance metrics commonly applied in contemporary stock assessment practices. Our results provide empirical evidence for the value of ecosystem-informed model-based approach, demonstrating that the inclusion of environmental and predator covariates has impact in model fit, predictive skill, and structural uncertainty.







\newpage

## Methodology

### Study Area

The study area covers the WAP, corresponding to FAO Subarea 48.1, one of the main regions for krill fishing activity. This area is also where CCAMLR is actively revising management approach at a finer spatial scale. To enhance the spatial resolution of our analysis of krill population dynamics, we used the five Management Units (MUs) (strata) proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest sector [@Dornam2021] (Fig. 1, S1). It is worth noting that these strata were originally developed for management purposes, without explicitly considering biological, ecological or population-specific criteria [@Watters2013; @Constable2023; @Zaldua2024a]. Nevertheless, their subdivision enables a finer-scale analysis of krill data, increasing spatial resolution and enabling detection of heterogeneity among various data sources—whether from fisheries or environmental variables—thus providing a more detailed understanding of krill population dynamics.


## Spatially Implicit Model

To represent spatial heterogeneity in krill dynamics, we implemented a spatially implicit modeling approach based on the *"areas-as-fleets"* methodology [@Waterhouse2014; @Punt2019; @Nielsen2021]. This approach treats distinct geographical regions as independent fleets, under the assumption that fishing and survey operations may have varying effects across space. Within this approach, each MU of the WAP is treated as a distinct fleet, allowing spatial differentiation in key parameters such as selectivity, catchability (*q*) and abundance indices. This is especially relevant for krill due to documented spatial variability and heterogeneity in both population processes and fishing effort [@Perry2019; @Perry2020; @Reiss2020; @Veytia2021; @SantaCruz2022]. By modeling each MU as a separate fleet, the assessment can estimate region-specific parameters that reflect this variability. Survey indices and biological compositions (e.g., length-frequency data) are assigned to their corresponding MU based on sampling location as defined by [@Dornam2021]. Similarly, fishery-dependent data—such as catch and size composition—are stratified by MU according to the reported fishing effort, allowing for regional variation in both abundance and observational processes. Given known difference in the population structure [@Kinzey2011; @Reiss2020; @Cutter2022], the selectivity is modeled separately for each MU to account for potential variations in size composition due to spatial heterogeneity or environmental conditions. This disaggregated structure allows the model to reflect small-scale spatial dynamics while maintaining analytical simplicity. The areas-as-fleets approach effectively incorporates regional variability in both population dynamics and observational processes without the added complexity of explicitly modeling movement [@Punt2019]. The model accounts for interannual variability in growth, maturity, and reproduction, assuming a single annual spawning pulse occurring between January and February [@Perry2019; @McBride2021]. Although MUs are treated as separate fleets, core population processes—such as recruitment and growth—are modeled collectively, representing the krill stock as a single interconnected population. This provides a more realistic representation of the krill population dynamics within Subarea 48.1.


## Length-to-Age Model Description

Like many invertebrates, krill present challenges for age determination, which complicates modeling population dynamics. Most modern stock assessments rely on age-structured modeling approaches to estimate marine population parameters [@Punt2013]. Although recent advancements have improved krill aging techniques [@Kilada2017], such methods have historically been unavailable for krill. Consequently, stock assessment for krill typically uses length data to approximate age structure through transition matrices. These matrices estimate the probability that individuals of a given length belong to specific age classes [@Laslett2004; @Lee2019; @Rudd2021; @Zhang2022; @Lee2024]. Given the scarcity—or absence—of direct age data in krill monitoring programs, length-based models offer a practical alternative to age-structured approaches. To date, all model-based krill assessments have used length-to-age frameworks [@Kinzey2015; @Kinzey2019; @Wang2021; @Green2023]. Integrated models can effectively capture the age structure by transforming length observations into population-level dynamics [@Punt2013; @Lee2024]. On the other hand, length data, which are cost-effective and readily available [@Chong2019; @Canales2021], provide valuable insights into krill population structure due to their correlation with age [@Thanassekos2014]. To perform this transformation, models often use Age-Length Keys (ALKs). An ALK is a probabilistic matrix that estimates the likelihood of individuals of a given length belonging to specific age classes. This method assumes both aged and measured individuals are random samples from the same population and should be applied within the same time period to avoid biases [@ICCA2003; @Punt2003; @Punt2013; @Lee2024]. The ALK matrix is typically defined as follows:


$$
P(A = a \mid L = l) = \frac{N_{a,l}}{\sum_{a} N_{a,l}}
$$
where \( P(A = a \mid L = l) \) is the probability that an individual of length \( l \) belongs to age \( a \),  \( N_{a,l} \) is the number of individuals of age \( a \) observed at length \( l \) and the denominator \( \sum_{a} N_{a,l} \) ensures that the probabilities sum to 1 for each length class. This matrix is then used in the model to allocate observed length compositions into age classes, typically through multiplication with a length-frequency vector:

$$
N_a = \sum_{l} P(A = a \mid L = l) \cdot N_l
$$

where \( N_l \) is the number of fish observed at length \( l \), and \( N_a \) is the estimated number at age. The resulting ALK is shown in S1 Fig 7. 

In the stock assessment catch-at-length model, growth is modeled using a von Bertalanffy growth function lenght-at-age parameterization with parameters L~inf~, k, and coefficient of variation (CV), as outlined in @Mardones2023. This growth relationship is integrated into the model alongside weight-at-length, which is assumed known and calculated using the parameters from @Maschette2020 for krill. The model assumed linear growth below a predefined settlement age of 1.5 years. The asymptotic length ($L_{\infty}$) was determined by the estimated length-at-age, indicating that the model used $L_{\infty}$ as an upper growth boundary. Exponential decay of growth beyond maximum age was disabled, ensuring that growth patterns followed the von Bertalanffy assumption without artificial constraints. Growth variability was not explicitly modeled, and the coefficient of variation (CV) was assumed to be a function of length-at-age. Maturity was modeled as an age-based logistic function, with the first age of maturity set at 2-3 years. Fecundity was estimated as a power function of body weight, where fecundity was computed as $Eggs = Wt \times (a + b \times Wt)$. The model used a direct parameterization without additional offsets between sexes. Regarding selectivity functions, it was modeled using a logistic formulation for the fishery fleet. The selectivity curve was defined by two key parameters: the size at 50% selectivity (inflection point) and the width of the selection range (95%). The fishery was characterized by multiple selectivity patterns, each defined by an inflection point and a selection range width. The estimated inflection points varied between 2.0 and 3.5 cm, while the selection range widths spanned between 2.0 and 4.0 cm depending of strata or "fleet". The survey selectivity was also modeled using a length-based logistic function. Inflection points ranged from 2.0 to 3.0 cm, with selection range widths consistently set at 1.0 cm. The phase assignment dictated whether parameters were freely estimated (positive phase values) or fixed (negative phase values). These differences in selectivity parameters are crucial for interpreting and modeling spatial heterogeneity among strata, as they reflect variability in size-dependent catchability across different regions and fleets (S1 Fig 6). The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm. assuming that the predator can capture a wider range of krill sizes. These parameters were estimated in later phases, allowing flexibility in their estimation. Across all selectivity functions, parameters with negative phase values were held fixed during the estimation process, ensuring model stability and preventing overparameterization in a initial step. The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt model curve like @Kinzey2019, with the steepness parameter fixed at 0.85 in phase -4 to ensure consistency with previous studies. This reflects a weak stock-recruitment relationship, as krill recruitment largely depends on environmental conditions rather than spawning biomass [@Dong2025; @Kinzey2019; @Perry2020]. The unfished recruitment ($R_0$) was estimated with a mean recruitment of 23 in log-scale, with a prior of 10. The parameter was freely estimated in phase 1. The standard deviation of recruitment deviations ($SR\_sigmaR$) was set at 1.2 and fixed in phase -4 to constrain model variability. The regime shift parameter ($SR\_regime$) was set at 0 and fixed, implying no consideration of long-term recruitment regime shifts. Recruitment deviations were estimated to account for interannual variability in recruitment dynamics. The model used the deviation vector approach ($do\_recdev = 1$), where recruitment was modeled as a function of spawning stock biomass plus a deviation term. The main recruitment deviation period spanned from 1998 to 2020, with forecast deviations applied in subsequent years. Early recruitment deviations prior to 1998 were included, with an early deviation phase of -4 to ensure these values remained fixed. Bias adjustment was implemented, with key adjustment years defined as 1995.7 as the last early year with no bias adjustment, 2000.0 as the first year of full bias adjustment, 2015.7 as the last year of full bias adjustment, and 2022.3 as the first recent year with no bias adjustment. To prevent overparameterization, recruitment deviations were constrained between -5 and 5. Cyclic patterns in recruitment were assumed each tree and five years [@Perry2019; @McBride2021; @Kinzey2019; @Thanassekos2014], and advanced recruitment options were enabled for model flexibility. Life history parameters—such as growth, weight-length relationship, natural mortality, and maturity—were used as priors to model the initial conditions of krill population dynamics, based on data from @Smith2023a, @Maschette2020, and @Kinzey2011. These parameters, along with initial selectivity assumptions, are summarized in Table \@ref(tab:parainit1). 




The statistical model to krill was configured using Stock Synthesis v 3.30.21 (SS3 hereafter) [@Methot2013]. SS3 is a structured stock assessment model, in the class of models called *"Integrated stock assessment model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment, and mortality processes, and observation, with expected values for different types of data. The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters. Model conditioning is available in the code repository section of Suppl. Mat 1. All analysis about modeling was executed in R-CRAN [@R-base] and diagnoses and visualization of the outputs was made was through *r4ss* [@Taylor2019] and *ss3diags* packages [@Winker2023]. 


## Environmental variable modelling

To identify the most influential environmental drivers of krill recruitment to be considered in stock evaluation processes, we examined the relationship between a recruitment index and three key environmental variables described previously as critical in shaping krill population dynamics: Sea Surface Temperature (SST), Sea Ice Concentration (SIC), and Chlorophyll-a concentration (Chla) [@Flores2012; @Morley2020; @Saba2014; @Veytia2021; @Walsh2020]. Table \@ref(tab:env) summarizes the key characteristics of the environmental datasets used in the analysis, including sea ice, sea surface temperature (SST), and chlorophyll a. These variables were obtained from different satellite platforms and reanalysis products, namely Nimbus-7, ERA5, and Copernicus. The table details spatial and temporal resolution, as well as the temporal coverage of each dataset.

```{r env, eval=TRUE}
env_features <- data.frame(
  Feature = c("Sea Ice", "SST", "Chlorophyll a"),
  Platform = c("Nimbus-7", "ERA5", "E.U. Copernicus Marine Service"),
  Spatial_Resolution = c("25 × 25 km", "0.25º × 0.25º", "0.25º × 0.25º"),
  Temporal_Resolution = c("Daily", "Monthly", "Daily-Monthly"),
  Time_Range = c("1978 - Present", "1991 - Present", "2000 - Present"),
  stringsAsFactors = FALSE
)

#to word
ft_env <- flextable(env_features) |>
  set_header_labels(
    Feature = "Feature",
    Platform = "Platform",
    Spatial_Resolution = "Spatial Resolution",
    Temporal_Resolution = "Temporal Resolution",
    Time_Range = "Time Range"
  ) |>
  set_caption("Summary of environmental variables used in correlation analysis") |>
  add_footer_lines("doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148") |>
  color(part = "header", color = "black") |>
  theme_booktabs() |>
  fontsize(size = 10) 

ft_1 <- autofit(ft_env)
ft_2 <- set_table_properties(ft_1, width = .5, layout = "autofit")
ft_2
```


All environmental data were processed at their original spatial resolution using the R packages *sf* and *ncdf4* [@Pebezma2018; @Pierce2021], incorporating spatial manipulation, rasterization, and coordinate transformations. We modelled the relation the environmental influence on krill recruitment with a serie of linear models, progressively incorporating the environmental variables to identify the one with the greatest explanatory power. The initial model considered only the effect of strata (spatial), while subsequent models sequentially included Chla, SST, SIC and interaction terms. This modeling framework allowed us to determine the relative contribution of each factor and their interactions on the recruitment process. All models and their results can be found in the Supplementary Material 2.



One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of population dynamics [@Marshall2019] and SS3 addresses these approaches. According to [@methot2020stock], Stock Synthesis (SS3) can estimate recruitment levels as a function of the spawning biomass from the previous season, incorporating the effects of an environmental time series and a log-bias adjustment. This integrated approach allows for a more realistic and dynamic estimation of recruitment by linking it to both biological and environmental drivers in this way;

$$
E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
$$
$R$ represents the variability of deviations, adding to the variance caused by environmental factors. $SpBio$ represent spawning biomass, $envdata$ is vector of time series of environmental variable. Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual $R$ should be reduced. However, the model does not automatically adjust for this. Based on the results of linear models, we will identify the environmental variable that exhibits the strongest statistical influence on krill recruitment. This variable will be selected for inclusion in the stock assessment model to capture the environmental contribution to recruitment variability. By integrating the most influential environmental driver, the model is expected to enhance the biological realism of recruitment estimates and improve the accuracy of population dynamics esttimates.


## Predator modelling 


Following @methot2020stock, natural mortality due to predation (*M2*) was incorporate through different types of observational data types, with each informing distinct aspects of predator-prey dynamics. In our krill assessment, we incorporated predator abundance from the three main penguin species in the Antarctic Peninsula: Adélie (*Pygoscelis adeliae*), Chinstrap (*Pygoscelis antarcticus*), and Gentoo (*Pygoscelis papua*)  from 399 colonies identified, located along the northern arc of the Antarctic Peninsula during the last two decades (2000–2020) obtained from the Mapping Application for Penguin Populations and Projected Dynamics ([MAPPPD](https://www.penguinmap.com/mapppd/sources/)). Predator abundance is processed as an index of predation "effort" allowing time-varying estimation of predation mortality rates in time. This follows the equation:

$$
M2_t = Q \cdot I_t
$$

where \( M2_t \) is the predation mortality in year *t*, \( I_t \) is the predator abundance index, and \( Q \) is a catchability coefficient. Additionally, length-frequency data of krill obtained from penguin diet samples from AMLR Program are used to characterize the size-selectivity of predation. These data inform predator-specific selectivity functions for *M2*, enabling the model to represent how predation varies with krill size. Total natural mortality (*M*) is thus partitioned as \(M = M_1 + \sum M_2 \), where \( M_1 \) is background mortality and \( \sum M_2 \) accounts for predator-specific contributions. Total mortality is expressed as \( Z = M_1 + M_2 + F \) combining natural and fishing-induced mortality. This dual approach—using predator abundance and diet composition—supports a temporally and size-structured representation of predation mortality, aligning with ecosystem-based assessment principles.


## Model Input Data: Fishery and Survey

Along with data on predators and environmental conditions, a fundamental input to the model was derived from fishery-dependent data and scientific monitoring programs. Fishery-dependent data were obtained from the Krill Fishery Monitoring Program under the CCAMLR SISO framework, including total reported catches, standardized CPUE (adjusted by vessel, area, and depth), and krill length-frequency distributions data were standardized using a fixed bin structure of 28 length intervals, each representing the lower edge of 0.2 cm width bins from 1.2 to 6.6 cm. Length data were assigned to partition code 0 (combined discard and retained) and sex code 0 (combined sexes), with a fixed sample size of 75 individuals per haul. This ensured consistency across years and strata and fleets (fishery, survey and predator), facilitating reliable temporal and spatial comparisons of size structure, covering the period from 1997 to 2020. Fishery-independent data originated from U.S. AMLR Program acoustic surveys included standarized biomass indices conducted between 1991 and 2011. Krill length compositions from AMLR surveys (provided by the Ecosystem Science Division of NOAA) was manipulate in the same structure that fishery-dependent data and handled according to stock assessment model templates requirements (Suppl. Mat. 2). All these data, together with the previously described ecosystem variables (environmental and predator-related), along with their temporal coverage and spatial scale, are summarized in Table \@ref(tab:data-sources) and illustrated in in S1, Fig. 4.


```{r data-sources, eval=TRUE}
data <- data.frame(
  Source = c(
    "Krill Fishery Monitoring Program (CCAMLR SISO)",
    "Krill Fishery Monitoring Program (CCAMLR SISO)",
    "Krill Fishery Monitoring Program (CCAMLR SISO)",
    "AMLR Scientific Research Program",
    "Satellite Remote Sensing",
    "Predator Monitoring (MAPPPD)",
    "Predator Diet Composition (AMLR)"
  ),
  Description = c(
    "Total reported catches",
    "Standardized CPUE (by vessel, area, and depth)",
    "Length-frequency distributions (0.2 cm bins)",
    "Biomass indices by strata from net sampling",
    "Monthly chlorophyll-a concentration from EU Copernicus Marine Service (interpolated; doi:10.48670/moi-00021, doi:10.48670/moi-00148)",
    "Abundance by colony-specific trends of penguin populations",
    "Length-frequency of krill in predator diets"
  ),
  TemporalScale = c(
    "1997–2020",
    "1997–2020",
    "2000–2020",
    "1991–2011 (based on cruise periods)",
    "1997–present",
    "2000s–present",
    "1977–2020 (subset from 2000 used in this analysis)"
  ),
  SpatialScale = c(
    "By strata",
    "By strata",
    "By strata",
    "By strata",
    "Subarea 48.1",
    "By colony",
    "By colony"
  )
)

# # to word
# ft_data <- flextable(data) %>%
#   set_caption("Table: Data sources used in the krill population dynamics modeling") %>%
#   theme_vanilla() %>%
#   fontsize(size = 9, part = "all") %>%
#   font(fontname = "Arial", part = "all") %>%
#   autofit() %>% 
#   set_table_properties(width = .5, layout = "autofit")
# 
# ft_data

# to PDF

kbl(data, format = "latex", 
    booktabs = TRUE, 
    caption = "Table: Data sources used in the krill population dynamics modeling") %>%
  kable_styling(latex_options = c("hold_position", "striped"), 
                font_size = 9, 
                full_width = FALSE) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(2, width = "5cm") %>%
  column_spec(3, width = "3cm") %>%
  column_spec(4, width = "3cm")



```

\newpage

## Scenarios

We evaluated a set of model configurations that incorporate variations in predator "fleet" and environmental effects, and the impacts in spawning biomass, recruitment among others. Spatial heterogeneity was considered in all scenarios. The methodology involves assessing the statistical performance of the model under these configurations. Specifically, `s1.1` serves as the reference model, incorporating only fishery and survey data while accounting for spatial heterogeneity (implicit spatial model). The subsequent scenarios build upon this configuration to evaluate the effects of additional ecological components. `s1.2` extends `s1.1` by incorporating predator influences, allowing the model to assess the impact of predation on krill population dynamics. `s1.3` modifies `s1.1` by introducing an environmental variable, capturing potential environmental drivers. Finally, `s1.4` integrates both predator and environmental influences into the `s1.1` framework, providing a more realistic assessment of ecosystem interactions and their role in krill population dynamics (Table \@ref(tab:scenario)).

```{r scenario, eval=TRUE}
scenario_table <- data.frame(
  Scenario = c("s1.1", "s1.2", "s1.3", "s1.4"),
  Description = c(
    "Spatial data without environmental and predator components",
    "\"s1.1\" with predator components",
    "\"s1.1\" with environmental variable",
    "\"s1.1\" with both predator fleet and environmental variable"
  ),
  stringsAsFactors = FALSE
)
ft_scenarios <- flextable(scenario_table) |>
  set_header_labels(
    Scenario = "Scenario",
    Description = "Description"
  ) |>
  set_caption("Scenarios used for modelling dynamics in krill") |>
  theme_booktabs() |>
   fontsize(size = 10) |>
  autofit() %>% 
  set_table_properties(width = 1, layout = "autofit") %>% 
  color(part = "header", color = "black") 
ft_scenarios
```


Regarding performance and scenario comparison, @Carvalho2017;  @Carvalho2021b; @Merino2022a and @Kell2021 propose a set of "good practices" for conducting a structured diagnosis to ensure the reliability and robustness of stock assessment analyses. This diagnostic process was applied to the krill stock assessment and involved several key steps proposed by previous research. First, convergence analysis was conducted to evaluate whether the estimation algorithm reached a stable solution. This included reviewing parameter trace plots, gradients, and the Hessian matrix to ensure that all parameters were well-estimated.  Next, model fit to data was assessed by comparing observed and predicted values across key data sources, including catch, survey indices, and length composition. The *p-values* for the residual runs tests were computed for each of the four joint length compositions that are common in all models. The *p-values* were then combined into a single test statistic using Fisher's method @@Merino2022a:

$$ 
\chi^2_{2k} = -2 \sum \log(p_i)
$$

where \( p_i \) is the p-value for Length \( i \) and \( k \) are the degrees of freedom of the four *p-values* from the joint length compositions. Likelihood profiling was also carried out to quantify parameter uncertainty and identify the influence of priors. Regarding model fit, Pearson residuals of predicted length distributions for krill were evaluated across four modeling scenarios, each incorporating different levels of ecosystem complexity. Pearson residuals are used to assess the goodness-of-fit by standardizing the differences between observed and predicted values, allowing for the detection of systematic patterns or biases in model predictions across size classes and sampling sources. Retrospective analysis and hindcast cross-validation were performed by sequentially removing five years of data regarding @HurtadoF2015. As part of the restrospective analysis, the Mohn parametrer was calculated as the mean relative error, defined as:

$$ 
\rho_M = \frac{1}{h} \sum_{y=1}^{h} \left( \frac{X_{T - y} - \hat{X}_{T - y}}{\hat{X}_{T - y}} \right)
$$

where \( X \) represents the quantities (such as spawning stock biomass (SSB) or fishing mortality (F)) for which \(\rho_M\) was calculated, and \( \hat{X} \) is the corresponding estimate from the reference model fitted to the full dataset. Year \( y \) was nominally considered as season one, \( T \) refers to the terminal assessment data year, and \( h \) denotes the number of retrospective peels, removing recent years of data ("peeling") and re-estimating the model, allowing for the calculation of retrospective bias (*Rho*) for key metrics such as Spawning Stock Biomass (SSB) and Fishing Mortality (F). To evaluate statistical differences in model performance, we analyzed Root Mean Squared Error (RMSE) values across four scenario configurations (`s1.1`, `s1.2`, `s1.3`, and `s1.4`). The RMSE values were extracted from the variance tuning diagnostics of each scenario and compiled into a data frame for statistical analysis. To test for overall differences among scenarios, a one-way ANOVA [@Sthle1989] was conducted, treating RMSE values as the dependent variable and scenario configurations as the categorical factor. Another quantitative statistical method used to evaluate model performance was the Akaike Information Criterion (AIC) [@Bozdogan1987]. This criterion balances model fit and complexity, facilitating comparisons across scenarios. Lower AIC values indicate more parsimonious models with better overall fit.

Table \@ref(tab:metrics) summarizes the diagnostic metrics used to evaluate model performance, including goodness-of-fit, prediction accuracy, information criteria, retrospective bias, and estimation stability. 

```{r metrics, eval=TRUE}
df_model_metrics <- data.frame(
  Metric = c(
    "Pearson Residuals",
    "Root Mean Square Error (RMSE)",
    "Mean Absolute Scaled Error (MASE)",
    "Akaike Information Criterion (AIC)",
    "Rho Parameter (B and SSB)",
    "R0 Deviation"
  ),
  Description = c(
    "Standardized residuals calculated as the difference between observed and expected values, divided by the standard deviation of the expected values. Useful for assessing model fit across data points.",
    "Measures the average magnitude of the residuals (prediction errors), giving higher weight to large errors. Lower values indicate better predictive accuracy.",
    "A scale-independent measure of prediction accuracy. It compares forecast errors to the errors from a naive baseline model. Values <1 indicate better-than-baseline performance.",
    "A relative measure of model quality that balances model fit and complexity. Lower AIC values indicate better models, penalizing overfitting.",
    "The autoregressive correlation coefficient from retrospective analysis, used to assess systematic bias in biomass (B) and spawning stock biomass (SSB) estimates across years.",
    "Measures the variability or uncertainty in the estimated unfished recruitment (R0). High deviations may indicate poor estimation stability or model overparameterization."
  ),
  Reference = c(
    NA,
    "@Carvalho2021b",
    "@Hyndman2006",
    "@Bozdogan1987",
    "@HurtadoF2015",
    "@Methot2013"
  ),
  stringsAsFactors = FALSE
)

# ft_stat <- flextable(df_model_metrics) |> 
#   set_header_labels(
#     Metric = "Metric",
#     Description = "Description",
#     Reference = "Reference"
#   ) |> 
#   set_caption("Summary of model diagnostic metrics used for krill stock assessment") |> 
#   #fontsize(size = 7) |> 
#   theme_booktabs() |> 
#   color(part = "header", color = "black")|>
#   set_table_properties(layout = "autofit")
# ft_stat


# to pdf

kbl(df_model_metrics, format = "latex", booktabs = TRUE,
    caption = "Summary of model diagnostic metrics used for krill stock assessment") |>
  kable_styling(latex_options = c("hold_position", "scale_down", "striped"), 
                font_size = 9) |>
  column_spec(1, width = "3.5cm", latex_valign = "m") |>
  column_spec(2, width = "9cm", latex_valign = "m") |>
  column_spec(3, width = "3.5cm", latex_valign = "m")

```
\pagebreak

## Results

### Environmental data 

Based on linear models performances and ecological relevance, Chla emerged as the most influential predictor, therefore, this variable was selected for incorporation into the stock assessment model as an environmental covariate (Table 2, S2).  Figure \@ref(fig:mapachl) presents a time series of Chl-a concentration (mg/m-3) in the waters around the Antarctic Peninsula from 2000 to 2020. Each panel represents a different year, showing spatial variations in Chl-a levels, which indicate phytoplankton biomass. The highest concentrations were observed along the coastal regions, while offshore areas exhibit lower Chl-a levels. 

\newpage

## Acknowledgments

This research was supported by the following funding sources: the INACH "Marine Protected Areas" Program (Grant No. 2409052), the ANID/Millennium Science Initiative Program (Grant No. ICN2021_002), the CCAMLR Scholarship Scheme (2023-2024), and the Doctorate Scholarship from CENTRO-IDEAL at the Universidad Austral de Chile. The authors extend their gratitude to the Secretariat of the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) for providing access to the krill fishery data, which was instrumental to this study.

## Supplementary Material 

-   Supplementary Material 1: analysis, formulas, and outputs of stock assessment of Antarctic krill in WAP can be found in this [link](https://mauromardones.github.io/SA_Krill/). 

-   Supplementary Material 2: Correlation analysis between krill recruitment and environmental variables can be found in this [link](https://mauromardones.github.io/Krill_recruit_env/) 


<!-- # Data Availability -->

<!-- Data will be made available on request. -->

\newpage


# The Second Appendix, for fun


