
```{r appendix, echo=FALSE}
torontodown::here_appendix()
```

# Manuscript 1 {#appendix1}

## "Disparate spatial heterogeneity in krill population dynamics across the Western Antarctic Peninsula"

This manuscript examines the spatial variability of Antarctic krill (*Euphausia superba*) population dynamics throughout the Western Antarctic Peninsula region. The study analyzes multi-year survey data to identify patterns of spatial heterogeneity in krill abundance, distribution, and demographic structure. Key findings reveal significant spatial clustering of krill populations driven by oceanographic features, sea ice dynamics, and regional environmental gradients.

**Publication Status:** Published in *Journal Name* (2024)  
**DOI:** [10.xxxx/journal.year.article](https://doi.org/10.xxxx/journal.year.article)  
**Authors:** [Author names]  
**Corresponding Chapter:** This work forms the foundation for [Chapter 3](#chap3) of this thesis.

**Keywords:** Antarctic krill, spatial heterogeneity, Western Antarctic Peninsula, population dynamics, Southern Ocean

\newpage

```{r convert-and-include, eval= FALSE, fig.align='center', out.width='100%'}
library(pdftools)
library(magick)
# Definir la ruta del PDF
pdf_path <- "~/DOCAS/LBSPR_Krill/LBSPR_Krill/LBSPR_Krill_MS.pdf"

# Verificar si el archivo existe
if(file.exists(pdf_path)) {
  
  # Convertir PDF a imágenes PNG con alta resolución
  images <- pdf_convert(pdf_path, format = "png", dpi = 100)
  
  # Mostrar información del documento
  pdf_info <- pdf_info(pdf_path)
  cat("Document:", basename(pdf_path), "\n")
  cat("Number of pages:", pdf_info$pages, "\n")
  cat("File size:", round(file.size(pdf_path)/1024/1024, 2), "MB\n\n")
  
  # Incluir todas las páginas
  knitr::include_graphics(images)
  
} else {
  cat("Error: PDF file not found at:", pdf_path, "\n")
  cat("Please check the file path and ensure the file exists.")
}
```


\newpage

# Manuscript 2 {#appendix2}

## "Stock assessment integrated Model in krill population in Western Antarctic Peninsula"

This manuscript examines the spatial variability of Antarctic krill (*Euphausia superba*) population dynamics throughout the Western Antarctic Peninsula region. The study analyzes multi-year survey data to identify patterns of spatial heterogeneity in krill abundance, distribution, and demographic structure. Key findings reveal significant spatial clustering of krill populations driven by oceanographic features, sea ice dynamics, and regional environmental gradients.

**Publication Status:** Published in *Journal Name* (2024)  
**DOI:** [10.xxxx/journal.year.article](https://doi.org/10.xxxx/journal.year.article)  
**Authors:** [Author names]  
**Corresponding Chapter:** This work forms the foundation for [Chapter 4](#chap4) of this thesis.

**Keywords:** Antarctic krill, spatial heterogeneity, Western Antarctic Peninsula, population dynamics, Southern Ocean

\newpage

```{r convert-and-include2, eval= FALSE, fig.align='center', out.width='100%'}


# Definir la ruta del PDF
pdf_path <- "~/DOCAS/WG/SAM_2025/WG_SAM_2025_11.pdf"

# Verificar si el archivo existe
if(file.exists(pdf_path)) {
  
  # Convertir PDF a imágenes PNG con alta resolución
  images <- pdf_convert(pdf_path, format = "png", dpi = 100)
  
  # Mostrar información del documento
  pdf_info <- pdf_info(pdf_path)
  cat("Document:", basename(pdf_path), "\n")
  cat("Number of pages:", pdf_info$pages, "\n")
  cat("File size:", round(file.size(pdf_path)/1024/1024, 2), "MB\n\n")
  
  # Incluir todas las páginas
  knitr::include_graphics(images)
  
} else {
  cat("Error: PDF file not found at:", pdf_path, "\n")
  cat("Please check the file path and ensure the file exists.")
}
```
